Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions

### 2017-08-17 draft.tex
Update codes, fix spelling

% if code changes, parameters are lost
Recording the execution environment is not enough though. The R$^2$ program uses a random seed but does not keep a trace of it except in the code. Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviouly be found. For instance, seed 1123581321 matches the results of R$^0$ with Python 2.7. This is not possible for a 100-step walk.}.

% 1. parameters and 2. code version with results
From this observation, we conclude two things. First, results files should always come alongside their context, i.e. an exhaustive list of the parameters used as well as a precise description of the execution environment, as the R$^3$ code does. Second, the version of the code itself should be recorded as part of the provenance data. It is common for results to have been generated by different versions of the code. Ideally, all results should originate from the same (and last) version of the code. But for long or expensive computations, this may not be feasible. In that case, the result files should contains the version of the code that was used to produce it. This information can be obtained from the version control software. This also allow, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed. In R$^3$, the code records the current git hash, and wether the repository holds uncommited changes when the computation starts.\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
import platform, datetime
import random
import git

def git_info(): # Retrieve the code version
    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
    return {'hash': repo.head.object.hexsha,
            'dirty': repo.is_dirty()}

def provenance(): # Get info about the execution environment
    return {'python'   : {'implementation': platform.python_implementation(),
                          'version'       : platform.python_version_tuple(),
                          'compiler'      : platform.python_compiler(),
                          'branch'        : platform.python_branch(),
                          'revision'      : platform.python_revision()},
            'platform' : platform.platform(),
            'git_info' : git_info(),
            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC

@@

Reusability of course directly benefits other researchers from your team and outside of it. The easier it is to use your code, the lower the threshold is for other to study, modify and extend it. Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage. A reproducible and reusable code offers a platform both \emph{verifiable} and easy-to-use, fostering the development of derivative works by other researchers on solid fundations. Those derivative works contribute to the impact of your original contribution.\\

Having more people examining and using your code also means that potential errors have a higher chance to be caught. If people start using your program, they will most likely report bugs or malfunctions they encounter. If you're lucky enough, they might even propose either bugfixes or improvements, hence improving the overall quality of your software. This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
import platform, datetime, json
import random
import git

def git_info():
    """Retrieve the current git hash, and if the repository has uncommited changes."""
    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
    return {'hash': repo.head.object.hexsha,
            'dirty': repo.is_dirty()}

def provenance():
    """Get info about the execution environment and the code."""
    return {'python'   : {'implementation': platform.python_implementation(),
                          'version'       : platform.python_version_tuple(),
                          'compiler'      : platform.python_compiler(),
                          'branch'        : platform.python_branch(),
                          'revision'      : platform.python_revision()},
            'platform' : platform.platform(),
            'git_info' : git_info(),
            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC

>>>>

% if code changes, parameters are lost
Recording the execution environment is not enough though. The R$^2$ program uses a random seed but does not keep a trace of it except in the code. Therefore, should the code change after the production of the results, someone provided with the result file will not be able to know which seed was used, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviously be found. For instance, seed 1123581321 matches the results of R$^0$ with Python 2.7. This is not possible for a 100-step walk.}.

% 1. parameters and 2. code version with results
From this observation, we conclude two things. First, results files should always come alongside their context, i.e. an exhaustive list of the parameters used as well as a precise description of the execution environment, as the R$^3$ code does. Second, the version of the code itself should be recorded as part of the provenance data. It is common for results to have been generated by different versions of the code. Ideally, all results should originate from the same (and last) version of the code. But for long or expensive computations, this may not be feasible. In that case, the result files should contain the version of the code that was used to produce it. This information can be obtained from the version control software. This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed. In R$^3$, the code records the current git hash, and whether the repository holds uncommited changes when the computation starts.\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 4:}} Re-runnable, repeatable, reproducible random walk (R$^3$)}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
import platform, datetime
import random
import pip
import git

def git_info(): # Retrieve the code version
    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
            'git_version': str(pip.git.Git().get_git_version())}

def provenance(): # Get info about the execution environment
    return {'python'   : {'implementation': platform.python_implementation(),
                          'version'       : platform.python_version_tuple(),
                          'compiler'      : platform.python_compiler(),
                          'branch'        : platform.python_branch(),
                          'revision'      : platform.python_revision()},
            'platform' : platform.platform(),
            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
            'git_info' : git_info(),
            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC

@@

Reusability of course directly benefits other researchers from your team and outside of it. The easier it is to use your code, the lower the threshold is for other to study, modify and extend it. Scientists constantly face the constraint of time: if a model is available, documented, and can be installed, run and understood all in a few hours, it will be preferred over another that would require weeks to reach the same stage. A reproducible and reusable code offers a platform both \emph{verifiable} and easy-to-use, fostering the development of derivative works by other researchers on solid foundations. Those derivative works contribute to the impact of your original contribution.\\

Having more people examining and using your code also means that potential errors have a higher chance to be caught. If people start using your program, they will most likely report bugs or malfunctions they encounter. If you're lucky enough, they might even propose either bug fixes or improvements, hence improving the overall quality of your software. This process contributes to the long-term reproducibility to the extent people continue to use and maintain the program.\\

@@

\noindent \begin{minipage}[c]{\linewidth}
\begin{code}{\textbf{\textsc{Listing 5:}} Re-runnable, repeatable, reproducible, reusable random walk (R$^4$)}
# Copyright (c) 2017 Nicolas P. Rougier and Fabien C. Y. Benureau
# Release under the BSD 2-clause license
# Tested with CPython 3.6.1 / macOS 10.12.4 / 64 bits architecture
import platform, datetime, json
import random
import pip
import git

def git_info():
    """Retrieve the current git hash, and if the repository has uncommited changes."""
    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
            'git_version': str(pip.git.Git().get_git_version())}

def provenance():
    """Get info about the execution environment and the code."""
    return {'python'   : {'implementation': platform.python_implementation(),
                          'version'       : platform.python_version_tuple(),
                          'compiler'      : platform.python_compiler(),
                          'branch'        : platform.python_branch(),
                          'revision'      : platform.python_revision()},
            'platform' : platform.platform(),
            'packages'  : list(pip.commands.freeze.freeze()), # list of installed packages
            'git_info' : git_info(),
            'timestamp': datetime.datetime.utcnow().isoformat()+'Z'}  # Z stands for UTC
